### 2021-02-25 tex/gitlinks.tex
try to fix environment

\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starrynight/blob/474eb00c5b87a26fb323129b3b73ea8e1b9dc20a/tex/figures/#1.py}{\codeicon}\,\,}
\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starrynight/blob/474eb00c5b87a26fb323129b3b73ea8e1b9dc20a/tex/figures/#1.gif}{\animicon}\,\,}
\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starrynight/blob/474eb00c5b87a26fb323129b3b73ea8e1b9dc20a/tex/proofs/#1.ipynb}{\raisebox{-0.1em}{\prooficon}}}
\newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starrynight/_build}{#1}}