### 2021-03-31 tex/gitlinks.tex
Committing automatically from GitHub Actions

\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starrynight/blob/5f79b24026133db207a88bee465e5e5080b64bcf/tex/figures/#1.py}{\codeicon}\,\,}
\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starrynight/blob/5f79b24026133db207a88bee465e5e5080b64bcf/tex/figures/#1.gif}{\animicon}\,\,}
\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starrynight/blob/5f79b24026133db207a88bee465e5e5080b64bcf/tex/proofs/#1.ipynb}{\raisebox{-0.1em}{\prooficon}}}
\newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starrynight/_build}{#1}}